---
title: "User Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| message: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
#| message: false

library(creelutils)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gt)

```

The `creelutils` package facilitates working with freshwater creel data by providing a toolkit that streamlines accessing data and performing common tasks in R.

# Accessing public data

Freshwater recreational creel data is made publicly available at Washington's official open data portal, [data.wa.gov](https:://data.wa.gov). Creel datasets are comprised of multiple elements (e.g., effort, interview, and catch data) that are grouped in space and time by a "Fishery Name" identifier. This identifier is an easy way to view a given creel's full dataset. 

For example, the *Skagit winter steelhead 2021* dataset contains effort counts, angler interviews, etc. conducted on sections of the Skagit and Sauk rivers from 2021-02-01 to 2021-04-15.

**By Fishery Name**

The function `fetch_dwg` downloads creel data from data.wa.gov by Fishery Name and is the primary route for accessing a complete dataset from a fishery.

```{r}
library(creelutils)

fishery_name <- "Skagit winter steelhead 2021"
dat <- fetch_dwg(fishery_name)

#ll stands for latitude and longitude
names(dat)

colnames(dat$effort)
```

**What if I don't know the Fishery Name?**

The function `fetch_fishery_names()`, which has no arguments, returns a list of all fishery names in the creel database.

```{r}
fetch_fishery_names() |> head(n = 10)
```

**What if I know part of the Fishery Name but not the specific syntax?**

The function `search_fishery_name()` can accept a vector of characters or numbers and returns the fishery names with those values.

```{r}
# by water body
search_fishery_name("Humptulips") |> head(n = 5)

# by run / species
search_fishery_name("winter steelhead") |> head(n = 5)

# by year
search_fishery_name("2025") |> head(n = 5)
```

## Accessing internal data

This section covers operations that require the appropriate credentials and permissions for WDFW's Postgres creel database. Selected 'views' from the relational database are mirrored on Washington's official open data portal, [data.wa.gov](https:://data.wa.gov), for public access.

### Connecting to database

An internal connection to the Postgres creel database can be made using the `establish_db_con` function. When called a RStudio UI prompt will open for the user to enter their password. This requires the proper user credentials and setup on local computer. If this is your first time using this method, contact the package maintainer for assistance. 

```{r}
# con <- establish_db_con()
```

