---
title: "creelutils User Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{creelutils User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
editor_options: 
  chunk_output_type: console
---

```{r}
#| include: false
#| message: false

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| include: false
#| message: false

library(creelutils)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gt)

```

The `creelutils` package was created to...

# Accessing public data

Freshwater recreational creel data is made publicly available at Washington's official open data portal, [data.wa.gov](https:://data.wa.gov). Creel datasets are comprised of multiple elements (e.g., effort, interview, and catch data) that are grouped in space and time by a "Fishery Name" identifier. This identifier is an easy way to view a given creel's full dataset. For example, the *Skagit winter steelhead 2021* dataset contains effort counts, angler interviews, etc. conducted on sections of the Skagit and Sauk rivers from 2021-02-01 to 2021-04-15.

**By Fishery Name**

The function `fetch_dwg` downloads creel data from data.wa.gov by Fishery Name.

```{r}
library(creelutils)

fishery_name <- "Skagit winter steelhead 2021"
dat <- fetch_dwg(fishery_name)

#ll stands for latitude and longitude
names(dat)

colnames(dat$effort)
```

**What if I don't know the Fishery Name?**

The function `fetch_fishery_names()`, which has no arguments, returns a list of all fishery names in the creel database.

```{r}
fetch_fishery_names() |> head(n = 10)
```

**What if I know part of the Fishery Name but not the specific syntax?**

The function `search_fishery_name()` can accept a vector of characters or numbers and returns the fishery names with those values.

```{r}
# by water body
search_fishery_name("Humptulips") |> head(n = 5)

# by run / species
search_fishery_name("winter steelhead") |> head(n = 5)

# by year
search_fishery_name("2025") |> head(n = 5)
```
